---
title: "AKSSAM: Real Data 2 - ElectricLoad"
author: "Nicolas Carrizosa Arias"
format: 
  html:
    toc: true        
    toc-depth: 4     
    toc-location: left  
    self-contained: true
---

# Functions

## Libraries

```{r library, warning = FALSE, message = FALSE}
# Required Packages
packages <- c(
  "qgam", "caret", "corpcor", "corrplot", "dplyr", "ggplot2", "grid",
  "gridExtra", "gridGraphics", "MASS", "Matrix", "patchwork", "pROC",
  "pracma", "splines", "splines2", "tidyverse", "aspline", "xtable",
  "cowplot", "scales", "mgcv", "pso"
)

# Install and/or load the packages
for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
  library(pkg, character.only = TRUE)
}
```

## AKKSAM

```{r AKSSAM, warning = FALSE, message=FALSE}
# AKSSAM
source("Required_Scripts/AKSSAM.R")
```

## OKPSO

```{r PSO, warning = FALSE, message=FALSE}
# PSO
source("Required_Scripts/ok_gam.R")
```


# Data

The considered dataset is `electric_load`, from the `R` package `opera`: Electricity forecasting data set provided by EDF R&D. It contains weekly measurements of the total electricity consumption in France from 1996 to 2009, together with several covariates, including temperature, industrial production indices (source: INSEE) and calendar information.

```{r}
# Load the dataset
load("Required_Scripts/electric_load.RData")

# Select the variables of interest
electric_load <- electric_load %>% na.omit
data <- electric_load %>% dplyr::select(Load1, NumWeek, Temp)
head(data)

## Pair Plot
data_no_dem <- data %>% scale(center = TRUE, scale = TRUE)
pairs(data_no_dem,
      pch = 19,
      main = "ElectricLoad",
      col = 'grey40',
      cex = 0.5)

## Corrplot
# Correlation matrix
cor_matrix <- cor(data_no_dem)
corrplot(cor_matrix, method = "color",
         addCoef.col = "black",                        
         number.cex = 0.7,                             
         tl.cex = 0.8,                                
         tl.col = "black",                             
         order = "hclust")
```


# Pipeline

## K-Fold function

```{r K-fold function}
kfold <- function(X, y, K){
  
  # Nº of instances and covariates
  n = length(y)
  m = dim(X)[2]
  
  # Identify the max and min covariates values' indexes
  ext_indexes = rep(0, 2*m) # Max of 2*m indexes
  grid <- matrix(0, ncol = m, nrow = 100) # Plotting grid
  for (t in 1:m){
    max_val = which.max(X[,t])
    min_val = which.min(X[,t])
    ext_indexes[2*(t-1) + 1] = max_val # Index for min obs of covariate t
    ext_indexes[2*(t-1) + 2] = min_val # Index for max obs of covariate t
    
    grid[,t] = seq(X[min_val,t], X[max_val,t], length.out = 100) # Grid for i-th covariate
  }
  # Eliminate repeated values and possible 0's
  ext_indexes = unique(ext_indexes[ext_indexes != 0])
  
  # Create the folds
  folds <- sample(rep(1:K, length.out = n)) 
  
  # Create a folds matrix which will have 1's if the data is included in the 
  # training sample for each fold (column)
  folds_matrix <- matrix(0, ncol = K, nrow = n)
  for (i in 1:K){
    # Add the extreme values
    training_idx <- unique(c(which(folds != i), ext_indexes)) 
    #Inidcate with 1 the training obs. indexes
    folds_matrix[training_idx, i] = 1
  }
  
  
  # Output
  return(list(folds = folds_matrix, # Folds matrix
              grid = grid           # Plotting grid
              ))
}
```


## Fitting function

OJO los nodos estásn puestos en ndx + bdeg + 1 para que, cuando no se dé ninguna seleccion de nodos, se tenga el mismo nº de elementos en la base para cada uno de los métodos

```{r}
fitting_function = function(X_train, X_test, y_train, y_test,
                            ndx, bdeg, lambda.init,
                            maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                            grid){
  
  # Nº of instances and covariates
  n = length(y_train)
  m = dim(X_train)[2]
  family = "gaussian"
  
  ## AKSSAM -------------------------------------------------------------------- 
  # Train the model
  time1 <- system.time({
    model1 <- AKSSAM(X_train, y_train, family, lambda.init, ndx, bdeg, 
                     maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, 
                     epsilon)
  })[3] %>% unname
  K_sel1 <- model1$K_sel
  alpha <- as.vector(model1$alpha.new)
  
  # Test
  Design_list = vector("list", m + 1)
  Design_list[[1]] = matrix(1, length(y_test), 1) 
  for (t in 1:m){Design_list[[t+1]] = my.bbase4(X_test[,t], K_sel1[[t]], bdeg[t])}
  B_star <- do.call(cbind, Design_list) 
  predictions1 <- B_star %*% alpha
  
  # Grid
  Grid_list = vector("list", m)
  grid_basis_length = rep(0,m)
  for (t in 1:m){     
    Grid_list[[t]] = my.bbase4(grid[,t], K_sel1[[t]], bdeg[t])
    grid_basis_length[t] = dim(Grid_list[[t]])[2]
  }
  grid1 <- grid
  for (t in 1:m){
    B = Grid_list[[t]]
    if (t == 1){
      index_0 = 2
      index_1 = 1 + grid_basis_length[1]
    }else{
      index_0 = 2 + sum(grid_basis_length[1:(t-1)])
      index_1 = index_0 + grid_basis_length[t] - 1
    }
    grid1[ ,t] <- B %*% alpha[index_0:index_1]
  }
  
  # Metrics
  MSE1 <- mean((y_test - predictions1)^2)
  MAE1 <- mean(abs(y_test - predictions1))
  EDF1 <- Size1 <- ncol(B_star)
  
  Old_Design_list = vector("list", m + 1)
  Old_Design_list[[1]] = matrix(1, length(y_train), 1)
  for (t in 1:m){Old_Design_list[[t+1]] = my.bbase4(X_train[,t], K_sel1[[t]], bdeg[t])}
  B_star_old <- do.call(cbind, Old_Design_list)
  residuals_train <- (y_train - B_star_old %*% alpha)
  sigma2_hat <- mean(residuals_train^2)
  loglik <- -0.5 * n * log(2 * pi * sigma2_hat) - 0.5 * sum(residuals_train^2) / sigma2_hat
  AIC1 <- -2 * loglik + 2 * EDF1
  BIC1 <- -2 * loglik + log(n) * EDF1
  
  
  ## PSO algorithm   -----------------------------------------------------------
  # Train the model
  data <- data.frame(y = y_train, as.data.frame(X_train))
  terms <- paste0("s(V", 1:m, ", bs = 'cr', k = ", ndx + bdeg + 1, ")")             
  formula <- as.formula(paste("y ~", paste(terms, collapse = " + ")))
  initial_gam_model <- mgcv::gam(formula, data = data)
  time2 <- system.time({
    result <- suppressMessages(
      fit_gam_pso(
        gam_model = initial_gam_model,
        data = data,
        n_knots = ndx[[1]] + bdeg[[1]] + 1,
        alpha = 1e-07,
        smoothing_method = "GCV.Cp",
        max_iterations = 100
      )
    )
  })[3] %>% unname
  model2 <- result$model
  K_sel2 <- result$knots %>% unname
  
  # Test
  test = as.data.frame(X_test)
  predictions2 <- as.vector(predict(model2, newdata = test))
    
  # Grid
  grid2 <- predict(model2, newdata = as.data.frame(grid), type = "terms")
  
  # Metrics
  MSE2 <- mean((y_test - predictions2)^2)
  MAE2 <- mean(abs(y_test - predictions2))
  EDF2 <- sum(model2$edf)
  Size2 <- length(model2$coefficients)
  
  residuals_train <- (y_train - predict(model2))
  sigma2_hat <- mean(residuals_train^2)
  loglik <- -0.5 * n * log(2 * pi * sigma2_hat) - 0.5 * sum(residuals_train^2) / sigma2_hat
  AIC2 <- -2 * loglik + 2 * EDF2
  BIC2 <- -2 * loglik + log(n) * EDF2
  
  
  ## P-splines   ---------------------------------------------------------------
  # Train the model
  terms <- paste0("s(V", 1:m, ", bs = 'ps', k = ", ndx + bdeg + 1, ")")             
  formula <- as.formula(paste("y ~", paste(terms, collapse = " + ")))
  time3 <- system.time({
    model3 <- mgcv::gam(formula, data = data, method = "REML")
  })[3] %>% unname
  
  # Test
  predictions3 <- as.vector(predict(model3, newdata = test))
    
  # Grid
  grid3 <- predict(model3, newdata = as.data.frame(grid), type = "terms")
  
  # Metrics
  MSE3 <- mean((y_test - predictions3)^2)
  MAE3 <- mean(abs(y_test - predictions3))
  EDF3 <- sum(model3$edf)
  Size3 <- length(model3$coefficients)
  
  residuals_train <- (y_train - predict(model3))
  sigma2_hat <- mean(residuals_train^2)
  loglik <- -0.5 * n * log(2 * pi * sigma2_hat) - 0.5 * sum(residuals_train^2) / sigma2_hat
  AIC3 <- -2 * loglik + 2 * EDF3
  BIC3 <- -2 * loglik + log(n) * EDF3
  
  
  ## Assignment   --------------------------------------------------------------
  return(list(
    # AKSSAM algorithm results
    AKSSAM = list(MSE = MSE1,
                  MAE = MAE1,
                  Grid = grid1,
                  K_sel = K_sel1,
                  EDF = EDF1,
                  Size = Size1,
                  AIC = AIC1,
                  BIC = BIC1,
                  Time = time1),
    # PSO algorithm results
    PSO = list(MSE = MSE2,
               MAE = MAE2,
               Grid = grid2,
               K_sel = K_sel2,
               EDF = EDF2,
               Size = Size2,
               AIC = AIC2,
               BIC = BIC2,
               Time = time2),
    # P-splines algorithm results
    PSplines = list(MSE = MSE3,
                    MAE = MAE3,
                    Grid = grid3,
                    EDF = EDF3,
                    Size = Size3,
                    AIC = AIC3,
                    BIC = BIC3,
                    Time = time3)
              ))
}
```


## Covariates Plot

```{r}
covariates_plot <- function(Results, grid, m, k){
  
  ## Miscellanea
  methods <- c('AKSSAM', 'PSO', 'P-splines')
  colors <- c("AKSSAM" = "#F8766D", "PSO" = "#7CAE00", "P-splines" = "#00BFC4")
        
  
  ## Calculate the mean estimated covariates' effects over the simulations
  # AKSSAM algorithm
  Grid_AKSSAM_matrices <- lapply(Results, function(x) x$AKSSAM$Grid)
  mean_Grid_AKSSAM <- Reduce("+", Grid_AKSSAM_matrices) / k
  temp <- lapply(Results, function(x) x$AKSSAM$K_sel)
  K_sel_AKSSAM <- lapply(1:m, function(i){
    unlist(lapply(1:k, function(t){
      v <- temp[[t]][[i]]
      return(v[5:(length(v)-4)])
    }))
  })
  
  # PSO algorithm
  Grid_PSO_matrices <- lapply(Results, function(x) x$PSO$Grid)
  mean_Grid_PSO <- Reduce("+", Grid_PSO_matrices) / k
  temp <- lapply(Results, function(x) x$PSO$K_sel) 
  K_sel_PSO <- lapply(1:m, function(i){
    unlist(lapply(1:k, function(t){
      v <- temp[[t]][[i]]
      return(v[2:(length(v)-1)])
    }))
  })
  
  # P-Splines
  Grid_Psp_matrices <- lapply(Results, function(x) x$PSplines$Grid)
  mean_Grid_Psp <- Reduce("+", Grid_Psp_matrices) / k
  
  
  # Combined grids
  mean_grids <- list(
    "AKSSAM" = mean_Grid_AKSSAM,
    "PSO" = mean_Grid_PSO,
    "P-splines" = mean_Grid_Psp
  )

  
  ## Marginalized covariates' effects
  plot_list <- vector('list', m)
  plot_list <- lapply(1:m, function(t) {
    
    if (t == 1){temp_title = "NumWeek - Week number"}
    if (t == 2){temp_title = "Temp - Temperature"}
    
    
    plot_df <- do.call(rbind, lapply(methods, function(method) {
      data.frame(
        x = grid[, t],
        y = mean_grids[[method]][, t],
        Method = method
      )
    }))
    plot_df$Method <- factor(plot_df$Method,
                         levels = c("P-splines", "PSO", "AKSSAM"))

    rug_df_AKSSAM <- data.frame(x = K_sel_AKSSAM[[t]])
    rug_df_PSO <- data.frame(x = K_sel_PSO[[t]])
    
    y_range <- range(plot_df$y, na.rm = TRUE)
    y_min <- y_range[1]
    y_max <- y_range[2]
    y_base <- y_min - 0.4 * (y_max - y_min)
    dy <- (y_max - y_min) * 0.07
    
    rug_df_AKSSAM$ymin <- y_base + dy
    rug_df_AKSSAM$ymax <- y_base + 2*dy
    
    rug_df_PSO$ymin <- y_base
    rug_df_PSO$ymax <- y_base + dy
    
    
    plot <- ggplot(plot_df, aes(x = x, y = y, color = Method)) +
      geom_line(linewidth = 1.1, alpha = 0.8, data = subset(plot_df, Method != "True effect")) +
      # 'Density' of selected knots
      geom_segment(data = rug_df_AKSSAM,
                 aes(x = x, xend = x, y = ymin, yend = ymax),
                 color = "#F8766D", linewidth = 2, alpha = 2/k) +
      geom_segment(data = rug_df_PSO,
                 aes(x = x, xend = x, y = ymin, yend = ymax),
                 color = "#7CAE00", linewidth = 2, alpha = 2/k) +
      scale_color_manual(values = colors) +
      theme_minimal() +
      scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
      labs(title = temp_title,
           color = 'Curve') +
      theme_light() + 
      theme(
        legend.position = "bottom",
        axis.title = element_blank(),
        panel.grid.minor = element_blank()
      )

      return(plot)
  })
  
  
  ## Combine the plots
  combined <- wrap_plots(plot_list, nrow = 1, guides = "collect") +
  plot_annotation(
    title = paste0("ElectricLoad dataset, ", k, "-fold"),
    theme = theme(
      plot.title = element_text(size = 15, face = "bold"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.background = element_rect(fill = alpha("white", 0.7), color = NA)
    )
  )
  
  ## Output
  return(combined)
}
```


## Error Plot

```{r}
error_plot <- function(Results){
  
  methods <- c("AKSSAM", "PSO", "PSplines")
  metrics <- c("MSE", "MAE", "EDF", "Size", "Time", "BIC")
  metric_labels <- c("MSE", "MAE", "EDF", "Basis Size", "Time", "BIC")
  
  all_data <- list()
  for (method in methods) {
    for (metric in metrics) {
      values <- sapply(Results, function(x) x[[method]][[metric]])
      temp_df <- data.frame(
        Value = values,
        Metric = metric,
        Method = method
      )
      all_data[[paste0(method, "_", metric)]] <- temp_df
    }
  }
  df <- bind_rows(all_data)
  df$Metric <- recode_factor(df$Metric,
    MSE = "MSE",
    MAE = "MAE",
    EDF = "EDF",
    Size = "Basis Size",
    Time = "Time",
    BIC = "BIC"
  )
  df$Value <- ifelse(df$Metric == "Time", log10(df$Value), df$Value)
  df$Metric <- recode_factor(df$Metric,
    MSE = "MSE",
    MAE = "MAE",
    EDF = "EDF",
    Size = "Basis Size",
    Time = "log10(Time)",
    BIC = "BIC"
  )
  metric_labels <- c("MSE", "MAE", "log10(Time)", "EDF", "Basis Size", "BIC")
  df$Metric <- factor(df$Metric, levels = metric_labels)
  
  df$Method <- recode(df$Method,
    "PSplines" = "P-splines",
    "ASplines" = "A-splines"
  )
  method_colors <- c(
    "AKSSAM"   = "#F8766D",
    "PSO"      = "#7CAE00",
    "P-splines" = "#00BFC4"
  )
  method_linetypes <- c(
    "AKSSAM"    = "solid",
    "PSO"       = "dashed",
    "P-splines" = "dotted"
  )
  
  # Boxplot of all performance metrics
  Performance_plot <- ggplot(df, aes(x = Method, y = Value, color = Method, linetype = Method)) +
    geom_boxplot(position = position_dodge(0.75), width = 0.6, fill = NA) +
    facet_wrap(~ Metric, scales = "free_y") +
    scale_color_manual(values = method_colors, 
                       name = "Algorithm",
                       labels = c("AKSSAM", "PSO", "P-splines", "A-splines")) +
    scale_linetype_manual(values = method_linetypes,
                          name = "Algorithm",
                          labels = c("AKSSAM", "PSO", "P-splines", "A-splines")) +
    labs(title = "Performance Metrics",
         x = "Method", y = "Value") +
    theme_light(base_size = 14) +
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      axis.text.x = element_text(size = 8),
    )

  
  ## Output
  return(Performance_plot)
}
```


## Wrapper

```{r}
wrapper_function = function(k, 
                            maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                            ndx, bdeg, lambda.init){
  
  # Set seed for reproducibility
  set.seed(13)
  m <- 2
  
  # Load the dataset
  load("Required_Scripts/electric_load.RData")

  # Select the variables of interest
  electric_load <- electric_load %>% na.omit
  data <- electric_load %>% dplyr::select(Load1, NumWeek, Temp)
  # Covariate matrix and response
  X <<- data %>% dplyr::select(-Load1) %>% as.matrix %>% scale(center = TRUE, scale = TRUE) %>% unname 
  attr(X, "scaled:center") <- NULL
  y <<- data %>% dplyr::select(Load1) %>% unlist %>% scale(center = TRUE, scale = TRUE) %>%  unname
  
  # List which contains the metrics for each simulation
  Results <- rep(list(
    list(
      AKSSAM = list(MSE = NA, MAE = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      PSO    = list(MSE = NA, MAE = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      PSplines = list(MSE = NA, MAE = NA, Grid = NA,
                      EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA)
    )
  ), k)
  
  
  ##  k-fold cross validation
  ll <- kfold(X, y, k)
  folds <- ll$folds
  grid <- ll$grid
  
  if (k == 1){
    # Train/Test equal
    X_train <- X_test <- X
    y_train <- y_test <- y
      
    # Fit the algorithms and store the results
    result <- fitting_function(X_train, X_test, y_train, y_test,
                                 ndx, bdeg, lambda.init,
                                 maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                                 grid)
    Results[[1]] <- result
  } else {
    # Fit the model for each split
    for (i in 1:k){
      # Obtain the Train/Test split
      train_idx <- which(folds[,i] == 1)
      X_train <- X[train_idx,] %>% as.matrix
      y_train <- y[train_idx]
      X_test <- X[-train_idx,] %>% as.matrix
      y_test <- y[-train_idx]
        
      # Fit the models
      result <- fitting_function(X_train, X_test, y_train, y_test,
                                   ndx, bdeg, lambda.init,
                                   maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                                   grid)
  
      # Store the result, increase index and restart try_attempts
      Results[[i]] <- result
      print(paste('Fold ', i, 'done.'))
    }
  }


  ## Output
  return(list(Results = Results,
              grid = grid))
  
}
```


## Processing Function

```{r}
process_function <- function(Results, grid, k){
  
  ## Miscelanea
  m <- 2
  
  ## Covariates estimated effect plot 
  cov_plot <- covariates_plot(Results, grid, m, k)
  
  
  ## Error metrics plot and values
  err_plot <- error_plot(Results)
  
  methods <- c("AKSSAM", "PSO", "PSplines")
  metrics <- c("MSE", "MAE", "AIC", "BIC", "EDF", "Size", "Time")
  
  err_metrics <- list()
  for (method in methods) {
    for (metric in metrics) {
      key_mean <- paste0("mean_", metric, "_", method)
      key_sd   <- paste0("sd_", metric, "_", method)
      
      values <- sapply(Results, function(x) {
        x[[method]][[metric]]
      })
      
      err_metrics[[key_mean]] <- mean(values, na.rm = TRUE)
      err_metrics[[key_sd]]   <- sd(values, na.rm = TRUE)
    }
  }
  
  ## Output
  return(list(covariates_plot = cov_plot,        # Covariates estimated effect plot 
              err_plot = err_plot,               # Error metrics plot
              err_metrics = err_metrics          # Error metrics
              ))
}
```


# GAM fit

```{r GAM fit}
ll <- wrapper_function(k = 5, 
                       maxiter1 = 100, maxiter2 = 100, maxiter3 = 100, 
                       tol1 = 1e-5, tol2 = 1e-5, tol3 = 1e-5, epsilon = 1e-5,
                       ndx = rep(40, 2), bdeg = rep(3, 2), lambda.init = rep(10, 2))
grid <- ll$grid
Results <- ll$Results
```

```{r GAM process, eval = FALSE}
results <- process_function(Results = ll$Results, grid = ll$grid, k = 5)
results$covariates_plot
results$err_plot
```

## Summary Table

### Full Table

```{r Summary Table, eval = FALSE}

# Define algorithms and metrics
algos <- c("AKSSAM", "PSO", "PSplines")
metrics <- c("MSE", "MAE", "EDF", "Size", "BIC", "Time")

# Initialize table with NA
cols <- c("Fold", unlist(lapply(algos, function(a) paste(metrics, a))))
results_table <- data.frame(matrix(NA, nrow = length(Results), ncol = length(cols)))
colnames(results_table) <- cols
results_table$Fold <- seq_along(Results)

# Fill the table
for (fold in seq_along(Results)) {
  res <- Results[[fold]]
  for (algo in algos) {
    for (metric in metrics) {
      colname <- paste(metric, algo)
      results_table[fold, colname] <- res[[algo]][[metric]]
    }
  }
}

# Display in HTML
knitr::kable(results_table, format = "html",
             caption = "Mean Results for the ElectricLoad dataset")

# Export to LaTeX
xtable::xtable(results_table, digits = 6)

```

# Save the RData

```{r Save the data}
# Checkpoint
save.image(file = "real_data2.RData")
```






