---
title: "Real Data 2 - PimaIndians - Binomial"
author: "Nicolas Carrizosa Arias"
format: 
  html:
    toc: true        
    toc-depth: 4     
    toc-location: left  
    self-contained: true
---

# Functions

## Libraries

```{r library, warning = FALSE, message = FALSE}

options(repos = c(CRAN="https://cloud.r-project.org"))

# Required Packages
if (!requireNamespace("qgam", quietly = TRUE)) install.packages("qgam")
library(qgam)

if (!requireNamespace("caret", quietly = TRUE)) install.packages("caret")
library(caret)

if (!requireNamespace("corpcor", quietly = TRUE)) install.packages("corpcor")
library(corpcor)

if (!requireNamespace("corrplot", quietly = TRUE)) install.packages("corrplot")
library(corrplot)

if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
library(dplyr)

if (!requireNamespace("ggplot2", quietly = TRUE)) install.packages("ggplot2")
library(ggplot2)

if (!requireNamespace("grid", quietly = TRUE)) install.packages("grid")
library(grid)

if (!requireNamespace("gridExtra", quietly = TRUE)) install.packages("gridExtra")
library(gridExtra)

if (!requireNamespace("gridGraphics", quietly = TRUE)) install.packages("gridGraphics")
library(gridGraphics)

if (!requireNamespace("MASS", quietly = TRUE)) install.packages("MASS")
library(MASS)

if (!requireNamespace("Matrix", quietly = TRUE)) install.packages("Matrix")
library(Matrix)

if (!requireNamespace("patchwork", quietly = TRUE)) install.packages("patchwork")
library(patchwork)

if (!requireNamespace("pROC", quietly = TRUE)) install.packages("pROC")
library(pROC)

if (!requireNamespace("pracma", quietly = TRUE)) install.packages("pracma")
library(pracma)

if (!requireNamespace("splines", quietly = TRUE)) install.packages("splines")
library(splines)

if (!requireNamespace("splines2", quietly = TRUE)) install.packages("splines2")
library(splines2)

if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
library(tidyverse)

if (!requireNamespace("aspline", quietly = TRUE)) install.packages("aspline")
library(aspline)

if (!requireNamespace("xtable", quietly = TRUE)) install.packages("xtable")
library(xtable)

if (!requireNamespace("cowplot", quietly = TRUE)) install.packages("cowplot")
library(cowplot)

if (!requireNamespace("scales", quietly = TRUE)) install.packages("scales")
library(scales)

if (!requireNamespace("mgcv", quietly = TRUE)) install.packages("mgcv")
library(mgcv)

if (!requireNamespace("pso", quietly = TRUE)) install.packages("pso")
library(pso)

if (!requireNamespace("aspline", quietly = TRUE)) install.packages("aspline")
library(aspline)
```

```{r library logistic, warning = FALSE, message = FALSE}
if (!requireNamespace("ROCR", quietly = TRUE)) install.packages("ROCR")
library(ROCR)
```


## AKKSAM

```{r AKSSAM, warning = FALSE, message=FALSE}
# AKSSAM
source("Required_Scripts/AKSSAM.R")
```


## GeDS

```{r GeDS, warning = FALSE, message=FALSE}
library(GeDS) 
```

# Data

The considered dataset is 'PimaIndiansDiabetes' form the 'mlbench' package.

```{r}
# Load the dataset
library(mlbench)
  data("PimaIndiansDiabetes")

  # Select the variables of interest
  PimaIndiansDiabetes <- PimaIndiansDiabetes %>% dplyr::select(mass, glucose, insulin, triceps, diabetes)
  data_no_dem <- PimaIndiansDiabetes %>% dplyr::select(-diabetes)

## Pair Plot
data_no_dem <- data_no_dem %>% scale(center = TRUE, scale = TRUE)
pairs(data_no_dem,
      pch = 19,
      main = "Smarket",
      col = 'grey40',
      cex = 0.5)

## Corrplot
# Correlation matrix
cor_matrix <- cor(data_no_dem)
corrplot(cor_matrix, method = "color",
         addCoef.col = "black",                        
         number.cex = 0.7,                             
         tl.cex = 0.8,                                
         tl.col = "black",                             
         order = "hclust")
```


# Pipeline

## K-Fold function

```{r K-fold function}
kfold <- function(X, y, K){
  
  # Nº of instances and covariates
  n = length(y)
  m = dim(X)[2]
  
  # Identify the max and min covariates values' indexes
  ext_indexes = rep(0, 2*m) # Max of 2*m indexes
  grid <- matrix(0, ncol = m, nrow = 100) # Plotting grid
  for (t in 1:m){
    max_val = which.max(X[,t])
    min_val = which.min(X[,t])
    ext_indexes[2*(t-1) + 1] = max_val # Index for min obs of covariate t
    ext_indexes[2*(t-1) + 2] = min_val # Index for max obs of covariate t
    
    grid[,t] = seq(X[min_val,t], X[max_val,t], length.out = 100) # Grid for i-th covariate
  }
  # Eliminate repeated values and possible 0's
  ext_indexes = unique(ext_indexes[ext_indexes != 0])
  
  # Create the folds
  folds <- sample(rep(1:K, length.out = n)) 
  
  # Create a folds matrix which will have 1's if the data is included in the 
  # training sample for each fold (column)
  folds_matrix <- matrix(0, ncol = K, nrow = n)
  for (i in 1:K){
    # Add the extreme values
    training_idx <- unique(c(which(folds != i), ext_indexes)) 
    #Inidcate with 1 the training obs. indexes
    folds_matrix[training_idx, i] = 1
  }
  
  
  # Output
  return(list(folds = folds_matrix, # Folds matrix
              grid = grid           # Plotting grid
              ))
}
```


## Fitting function

OJO los nodos estásn puestos en ndx + bdeg + 1 para que, cuando no se dé ninguna seleccion de nodos, se tenga el mismo nº de elementos en la base para cada uno de los métodos

```{r}
fitting_function = function(X_train, X_test, y_train, y_test,
                            ndx, bdeg, lambda.init,
                            maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                            grid){
  
  # Nº of instances and covariates
  n <- length(y_train)
  m <- dim(X_train)[2]
  family <- "binomial"
  
  
  ## AKSSAM -------------------------------------------------------------------- 
  # Train the model
  time1 <- system.time({
    model1 <- AKSSAM(X_train, y_train, family, lambda.init, ndx, bdeg, 
                     maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, 
                     epsilon)
  })[3] %>% unname
  K_sel1 <- model1$K_sel
  alpha <- as.vector(model1$alpha.new)
  
  # Test
  Design_list = vector("list", m + 1)
  Design_list[[1]] = matrix(1, length(y_test), 1) 
  for (t in 1:m){Design_list[[t+1]] = my.bbase4(X_test[,t], K_sel1[[t]], bdeg[t])}
  B_star <- do.call(cbind, Design_list) 
  eta1 <- B_star %*% alpha
  prob1 <- exp(eta1)/(1 + exp(eta1))
  
  # Grid
  Grid_list = vector("list", m)
  grid_basis_length = rep(0,m)
  for (t in 1:m){     
    Grid_list[[t]] = my.bbase4(grid[,t], K_sel1[[t]], bdeg[t])
    grid_basis_length[t] = dim(Grid_list[[t]])[2]
  }
  grid1 <- grid
  for (t in 1:m){
    B = Grid_list[[t]]
    if (t == 1){
      index_0 = 2
      index_1 = 1 + grid_basis_length[1]
    }else{
      index_0 = 2 + sum(grid_basis_length[1:(t-1)])
      index_1 = index_0 + grid_basis_length[t] - 1
    }
    grid1[ ,t] <- B %*% alpha[index_0:index_1]
  }
  
  # Metrics
  predictions1 <- ifelse(prob1 >= 0.5, 1, 0)
  ACC1 <- sum(y_test == predictions1) / length(y_test)
  AUC1 <- auc(roc(as.vector(as.numeric(y_test)), as.vector(as.numeric(prob1)), quiet = TRUE))[1]
  
  EDF1 <- Size1 <- ncol(B_star)
  
  Old_Design_list = vector("list", m + 1)
  Old_Design_list[[1]] = matrix(1, length(y_train), 1)
  for (t in 1:m){Old_Design_list[[t+1]] = my.bbase4(X_train[,t], K_sel1[[t]], bdeg[t])}
  B_star_old <- do.call(cbind, Old_Design_list)
  eta_old1 <- B_star_old %*% alpha
  prob_old1 <- exp(eta_old1)/(1 + exp(eta_old1))
  
  loglik <- sum(y_train * log(prob_old1) + (1 - y_train) * log(1 - prob_old1))
  AIC1 <- -2 * loglik + 2 * EDF1
  BIC1 <- -2 * loglik + log(n) * EDF1
  
  
  ## P-splines   ---------------------------------------------------------------
  # Train the model
  data <- data.frame(y = y_train, as.data.frame(X_train))
  test = as.data.frame(X_test)
  terms <- paste0("s(V", 1:m, ", bs = 'ps', k = ", ndx + bdeg + 1, ")")             
  formula <- as.formula(paste("y ~", paste(terms, collapse = " + ")))
  time3 <- system.time({
    model3 <- mgcv::gam(formula, data = data, family = binomial(link = "logit"), method = "REML")
  })[3] %>% unname
  
  # Test
  prob3 <- as.vector(predict(model3, newdata = test , type = "response"))
    
  # Grid
  grid3 <- predict(model3, newdata = as.data.frame(grid), type = "terms")
  
  # Metrics
  predictions3 <- ifelse(prob3 >= 0.5, 1, 0)
  ACC3 <- sum(y_test == predictions3) / length(y_test)
  AUC3 <- auc(roc(as.vector(as.numeric(y_test)), as.vector(as.numeric(prob3)), quiet = TRUE))[1]
  
  
  EDF3 <- sum(model3$edf)
  Size3 <- length(model3$coefficients)
  
  prob_old3 <- as.vector(predict(model3, type = "response"))
  loglik <- sum(y_train * log(prob_old3) + (1 - y_train) * log(1 - prob_old3))
  AIC3 <- -2 * loglik + 2 * EDF3
  BIC3 <- -2 * loglik + log(n) * EDF3
  
  ## GeDS ----------------------------------------------------------------------
  # Train the model
  terms_GeDS   <- paste0("f(V", 1:m, ")")
  formula_GeDS <- as.formula(paste("y ~", paste(terms_GeDS, collapse = " + ")))
  time4 <- system.time({
    model4 <- GeDS::NGeDSgam(formula_GeDS, family = binomial(link = "logit"), data = data, internal_knots = ndx[[1]] - 1)
  })[3] %>% unname
   K_sel4 <- model4$internal_knots$cubic.int.knots %>% unname
  
  # Test
  prob4 <- as.vector(predict(model4, n = 4, newdata = test , type = "response"))
  
  # grid
  grid4 <- do.call(cbind, lapply(terms_GeDS, function(term) {
    matrix(
      predict(model4, n = 4, newdata = as.data.frame(grid), base_learner = term),
      ncol = 1, dimnames = list(NULL, term)
    )
  }))
  
  # Metrics
  predictions4 <- ifelse(prob4 >= 0.5, 1, 0)
  ACC4 <- sum(y_test == predictions4) / length(y_test)
  AUC4 <- auc(roc(as.vector(as.numeric(y_test)), as.vector(as.numeric(prob4)), quiet = TRUE))[1]
  
  K_len4 <- lapply(knots(model4, n = 4L, options ="all"), length) %>% unname
  
  EDF4 <- Size4 <- sum(sapply(knots(model4, n = 4L, options ="all"), length) %>% unname - bdeg - 1) + 1 # intercept
  
  prob_old4 <- as.vector(predict(model4, n = 4, type = "response"))
  loglik4 <- sum(y_train * log(prob_old4) + (1 - y_train) * log(1 - prob_old4))
  AIC4 <- -2 * loglik4 + 2 * EDF4
  BIC4 <- -2 * loglik4 + log(n) * EDF4
  
  
  ## Assignment   --------------------------------------------------------------
  return(list(
    # AKSSAM algorithm results
    AKSSAM = list(ACC = ACC1,
                  AUC = AUC1,
                  Grid = grid1,
                  K_sel = K_sel1,
                  EDF = EDF1,
                  Size = Size1,
                  AIC = AIC1,
                  BIC = BIC1,
                  Time = time1),
    # P-splines algorithm results
    PSplines = list(ACC = ACC3,
                    AUC = AUC3,
                    Grid = grid3,
                    EDF = EDF3,
                    Size = Size3,
                    AIC = AIC3,
                    BIC = BIC3,
                    Time = time3),
    # GeDS algorithm results
    GeDS = list(ACC = ACC4,
                    AUC = AUC4,
                    Grid = grid4,
                    K_sel = K_sel4,
                    EDF = EDF4,
                    Size = Size4,
                    AIC = AIC4,
                    BIC = BIC4,
                    Time = time4)
              ))
}
```


## Covariates Plot

```{r}
covariates_plot <- function(Results, grid, m, k){
  
  ## Miscellanea
  methods <- c('AKSSAM', 'P-splines', 'GeDS')
  colors <- c("AKSSAM" = "#F8766D", "P-splines" = "#00BFC4", 'GeDS' = '#EE9A49')
        
  
  ## Calculate the mean estimated covariates' effects over the simulations
  # AKSSAM algorithm
  Grid_AKSSAM_matrices <- lapply(Results, function(x) x$AKSSAM$Grid)
  mean_Grid_AKSSAM <- Reduce("+", Grid_AKSSAM_matrices) / k
  temp <- lapply(Results, function(x) x$AKSSAM$K_sel)
  K_sel_AKSSAM <- lapply(1:m, function(i){
    unlist(lapply(1:k, function(t){
      v <- temp[[t]][[i]]
      return(v[5:(length(v)-4)])
    }))
  })
  
  # P-Splines
  Grid_Psp_matrices <- lapply(Results, function(x) x$PSplines$Grid)
  mean_Grid_Psp <- Reduce("+", Grid_Psp_matrices) / k
  
  
    # GeDS algorithm
  Grid_GeDS_matrices <- lapply(Results, function(x) x$GeDS$Grid)
  mean_Grid_GeDS <- Reduce("+", Grid_GeDS_matrices) / k
  mean_Grid_GeDS <- apply(mean_Grid_GeDS, 2, function(x) x - mean(x)) ## Possible misspecification
  temp <- lapply(Results, function(x) x$GeDS$K_sel)
  K_sel_GeDS <- lapply(1:m, function(i){
    unlist(lapply(1:k, function(t){
      v <- temp[[t]][[i]]
      return(v)
    }))
  })
  
  # Combined grids
  mean_grids <- list(
    "AKSSAM" = mean_Grid_AKSSAM,
    "P-splines" = mean_Grid_Psp,
    'GeDS' = mean_Grid_GeDS
  )

  
  ## Marginalized covariates' effects
  plot_list <- vector('list', m)
  plot_list <- lapply(1:m, function(t) {
    
    if (t == 1){temp_label = "mass"}
    if (t == 2){temp_label = "age"}
    
    
    plot_df <- do.call(rbind, lapply(methods, function(method) {
      data.frame(
        x = grid[, t],
        y = mean_grids[[method]][, t],
        Method = method
      )
    }))
    plot_df$Method <- factor(plot_df$Method,
                         levels = c("P-splines", "AKSSAM", 'GeDS'))

    y_range <- range(plot_df$y, na.rm = TRUE)
    y_min <- y_range[1]
    y_max <- y_range[2]
    y_base <- y_min - 0.4 * (y_max - y_min)
    dy <- (y_max - y_min) * 0.07
    
    rug_df_AKSSAM <- if (length(K_sel_AKSSAM[[t]]) > 0) {
        data.frame(x = K_sel_AKSSAM[[t]], 
                   ymin = y_base + dy, 
                   ymax = y_base + 2*dy)
      } else {
        data.frame(x = numeric(0), ymin = numeric(0), ymax = numeric(0))
      }
      
      rug_df_GeDS <- if (length(K_sel_GeDS[[t]]) > 0) {
        data.frame(x = K_sel_GeDS[[t]], 
                   ymin = y_base, 
                   ymax = y_base + dy)
      } else {
        data.frame(x = numeric(0), ymin = numeric(0), ymax = numeric(0))
      }
  
    
    
    plot <- ggplot(plot_df, aes(x = x, y = y, color = Method)) +
      geom_line(linewidth = 1.1, alpha = 0.8, data = subset(plot_df, Method != "True effect")) +
      # 'Density' of selected knots
      geom_segment(data = rug_df_AKSSAM,
                   aes(x = x, xend = x, y = ymin, yend = ymax),
                   inherit.aes = FALSE,
                   color = "#F8766D", linewidth = 2, alpha = 2/k) +
      geom_segment(data = rug_df_GeDS,
                   aes(x = x, xend = x, y = ymin, yend = ymax),
                   inherit.aes = FALSE,
                   color = "#EE9A49", linewidth = 2, alpha = 2/k) +
      scale_color_manual(values = colors) +
      theme_minimal() +
      scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
      labs(x = temp_label,
           y = expression(eta[diabetes]),
           color = 'Curve') +
      theme_light() + 
      theme(
        legend.position = "bottom",
        panel.grid.minor = element_blank()
      )

      return(plot)
  })
  
  
  ## Combine the plots
  combined <- wrap_plots(plot_list, nrow = 1, guides = "collect") +
  plot_annotation(
    theme = theme(
      plot.title = element_text(size = 15, face = "bold"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.background = element_rect(fill = alpha("white", 0.7), color = NA)
    )
  )
  
  ## Output
  return(combined)
}
```


## Error Plot

```{r}
error_plot <- function(Results){
  
  methods <- c("AKSSAM", "PSplines", 'GeDS')
  metrics <- c("ACC", "AUC", "EDF", "Size", "Time", "BIC")
  metric_labels <- c("ACC", "AUC", "EDF", "Basis Size", "log10(Time)", "BIC")
  
  all_data <- list()
  for (method in methods) {
    for (metric in metrics) {
      values <- sapply(Results, function(x) x[[method]][[metric]])
      temp_df <- data.frame(
        Value = values,
        Metric = metric,
        Method = method
      )
      all_data[[paste0(method, "_", metric)]] <- temp_df
    }
  }
  df <- bind_rows(all_data)
  df$Value <- ifelse(df$Metric == "Time", log10(df$Value), df$Value)
  df$Metric <- recode_factor(df$Metric,
    ACC = "ACC",
    AUC = "AUC",
    EDF = "EDF",
    Size = "Basis Size",
    Time = "log10(Time)",
    BIC = "BIC"
  )
  df$Metric <- factor(df$Metric, levels = metric_labels)
  df$Metric <- factor(df$Metric, levels = c("ACC", "AUC", "log10(Time)", "EDF", "Basis Size", "BIC"))
  
  df$Method <- dplyr::recode(df$Method,
    "PSplines" = "P-splines"
  )
  method_colors <- c(
    "AKSSAM"   = "#F8766D",
    "P-splines" = "#00BFC4",
    'GeDS' = '#EE9A49'
  )
  
  # Boxplot of all performance metrics
  Performance_plot <- ggplot(df, aes(x = Method, y = Value, color = Method)) +
    geom_boxplot(position = position_dodge(0.75), width = 0.6, fill = NA) +
    facet_wrap(~ Metric, scales = "free_y") +
    scale_color_manual(values = method_colors) +
    labs(title = "Performance Metrics",
         x = "Method", y = "Value", color = "Algorithm") +
    theme_light(base_size = 14) +
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      axis.text.x = element_text(size = 8),
    )

  
  ## Output
  return(Performance_plot)
}
```


## Wrapper

```{r}
wrapper = function(k, 
                            maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                            ndx, bdeg, lambda.init){
  
  # Set seed for reproducibility
  set.seed(12)
  
  # Load the dataset
  data("PimaIndiansDiabetes")

  # Select the variables of interest
  PimaIndiansDiabetes <- PimaIndiansDiabetes %>% dplyr::select(mass, age, diabetes) #dplyr::select(mass, insulin, age, diabetes)
  data <- PimaIndiansDiabetes %>% na.omit
  # Covariate matrix and response
  X <<- data %>% dplyr::select(-diabetes) %>% as.matrix %>% scale(center = TRUE, scale = TRUE) %>% unname 
  attr(X, "scaled:center") <- NULL
  attr(X, "scaled:scale") <- NULL
  m <<- dim(X)[2]
  y <<- data %>% dplyr::select(diabetes) %>% dplyr::mutate(diabetes = dplyr::if_else(diabetes == "pos", 1L, 0L)) %>% unname %>% unlist

  # List which contains the metrics for each simulation
  Results <- rep(list(
    list(
      AKSSAM = list(ACC = NA, AUC = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      PSplines = list(ACC = NA, AUC = NA, Grid = NA,
                      EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      GeDS = list(ACC = NA, AUC = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA)
    )
  ), k)
  
  
  ##  k-fold cross validation
  ll <- kfold(X, y, k)
  folds <- ll$folds
  grid <- ll$grid
  
  if (k == 1){
    # Train/Test equal
    X_train <- X_test <- X
    y_train <- y_test <- y
      
    # Fit the algorithms and store the results
    result <- purrr::quietly(fitting_function)(
        X_train = X_train,
        X_test = X_test,
        y_train = y_train,
        y_test = y_test,
        ndx = ndx,
        bdeg = bdeg,
        lambda.init = lambda.init,
        maxiter1 = maxiter1,
        maxiter2 = maxiter2,
        maxiter3 = maxiter3,
        tol1 = tol1,
        tol2 = tol2,
        tol3 = tol3,
        epsilon = epsilon,
        grid = grid
      )$result
    
    Results[[1]] <- result
  } else {
    # Fit the model for each split
    for (i in 1:k){
      # Obtain the Train/Test split
      train_idx <- which(folds[,i] == 1)
      X_train <- X[train_idx,] %>% as.matrix
      y_train <- y[train_idx]
      X_test <- X[-train_idx,] %>% as.matrix
      y_test <- y[-train_idx]
      n <- length(y_train)
      
      current_env <- sys.frame(sys.nframe())
      list2env(as.list(current_env), envir = globalenv())
      
      # Fit the models
      result <- purrr::quietly(fitting_function)(
        X_train = X_train,
        X_test = X_test,
        y_train = y_train,
        y_test = y_test,
        ndx = ndx,
        bdeg = bdeg,
        lambda.init = lambda.init,
        maxiter1 = maxiter1,
        maxiter2 = maxiter2,
        maxiter3 = maxiter3,
        tol1 = tol1,
        tol2 = tol2,
        tol3 = tol3,
        epsilon = epsilon,
        grid = grid
      )$result
  
      # Store the result, increase index and restart try_attempts
      Results[[i]] <- result
      print(paste('Fold ', i, 'done.'))
    }
  }


  ## Output
  return(list(Results = Results,
              grid = grid))
  
}
```


## Processing Function

```{r}
process<- function(Results, grid, k){
  
  ## Miscelanea
  m <- 2
  
  ## Covariates estimated effect plot 
  cov_plot <- covariates_plot(Results, grid, m, k)
  
  
  ## Error metrics plot and values
  err_plot <- error_plot(Results)
  
  methods <- c("AKSSAM", "PSplines", 'GeDS')
  metrics <- c("ACC", "AUC", "AIC", "BIC", "EDF", "Size", "Time")
  
  err_metrics <- list()
  for (method in methods) {
    for (metric in metrics) {
      key_mean <- paste0("mean_", metric, "_", method)
      key_sd   <- paste0("sd_", metric, "_", method)
      
      values <- sapply(Results, function(x) {
        x[[method]][[metric]]
      })
      
      err_metrics[[key_mean]] <- mean(values, na.rm = TRUE)
      err_metrics[[key_sd]]   <- sd(values, na.rm = TRUE)
    }
  }
  
  ## Output
  return(list(covariates_plot = cov_plot,        # Covariates estimated effect plot 
              err_plot = err_plot,               # Error metrics plot
              err_metrics = err_metrics          # Error metrics
              ))
}
```



# GAM fit

```{r GAM fit, eval = FALSE}
ll <- wrapper(k = 5, 
              maxiter1 = 100, maxiter2 = 100, maxiter3 = 100, 
              tol1 = 1e-5, tol2 = 1e-5, tol3 = 1e-5, epsilon = 1e-5,
              ndx = rep(40, 3), bdeg = rep(3, 3), lambda.init = rep(1, 3))
grid <- ll$grid
Results <- ll$Results
```

# Load Data

```{r load, eval = FALSE}
load("real_data2.RData")
```

# Process Results

```{r GAM process, eval = TRUE}
results <- process(Results = Results, grid = grid, k = 5)
results$err_plot
results$covariates_plot
```



## Summary Table

### Full Table

```{r Summary Table, eval = TRUE}

# Define algorithms and metrics
algos <- c("AKSSAM", "GeDS", "PSplines")
metrics <- c("ACC", "AUC", "EDF", "Size", "BIC", "Time")

# Initialize table with NA
cols <- c("Fold", unlist(lapply(algos, function(a) paste(metrics, a))))
results_table <- data.frame(matrix(NA, nrow = length(Results), ncol = length(cols)))
colnames(results_table) <- cols
results_table$Fold <- seq_along(Results)

# Fill the table
for (fold in seq_along(Results)) {
  res <- Results[[fold]]
  for (algo in algos) {
    for (metric in metrics) {
      colname <- paste(metric, algo)
      results_table[fold, colname] <- res[[algo]][[metric]]
    }
  }
}

# Display in HTML
knitr::kable(results_table, format = "html",
             caption = "Mean Results for the PimaIndians dataset")

# Export to LaTeX
xtable::xtable(results_table, digits = 6)

```

# Save the RData

```{r Checkpoint, eval = FALSE}
# save.image(file = "real_data2.RData")
```






