---
title: "Simulation 2 - Poisson"
author: "Nicolas Carrizosa Arias"
format: 
  html:
    toc: true        
    toc-depth: 4     
    toc-location: left  
    self-contained: true
---

# Functions

## Libraries

```{r library, warning = FALSE, message=FALSE}

options(repos = c(CRAN="https://cloud.r-project.org"))

# Required Packages
if (!requireNamespace("qgam", quietly = TRUE)) install.packages("qgam")
library(qgam)

if (!requireNamespace("caret", quietly = TRUE)) install.packages("caret")
library(caret)

if (!requireNamespace("corpcor", quietly = TRUE)) install.packages("corpcor")
library(corpcor)

if (!requireNamespace("corrplot", quietly = TRUE)) install.packages("corrplot")
library(corrplot)

if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
library(dplyr)

if (!requireNamespace("ggplot2", quietly = TRUE)) install.packages("ggplot2")
library(ggplot2)

if (!requireNamespace("grid", quietly = TRUE)) install.packages("grid")
library(grid)

if (!requireNamespace("gridExtra", quietly = TRUE)) install.packages("gridExtra")
library(gridExtra)

if (!requireNamespace("gridGraphics", quietly = TRUE)) install.packages("gridGraphics")
library(gridGraphics)

if (!requireNamespace("MASS", quietly = TRUE)) install.packages("MASS")
library(MASS)

if (!requireNamespace("Matrix", quietly = TRUE)) install.packages("Matrix")
library(Matrix)

if (!requireNamespace("patchwork", quietly = TRUE)) install.packages("patchwork")
library(patchwork)

if (!requireNamespace("pROC", quietly = TRUE)) install.packages("pROC")
library(pROC)

if (!requireNamespace("pracma", quietly = TRUE)) install.packages("pracma")
library(pracma)

if (!requireNamespace("splines", quietly = TRUE)) install.packages("splines")
library(splines)

if (!requireNamespace("splines2", quietly = TRUE)) install.packages("splines2")
library(splines2)

if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
library(tidyverse)

if (!requireNamespace("aspline", quietly = TRUE)) install.packages("aspline")
library(aspline)

if (!requireNamespace("xtable", quietly = TRUE)) install.packages("xtable")
library(xtable)

if (!requireNamespace("cowplot", quietly = TRUE)) install.packages("cowplot")
library(cowplot)

if (!requireNamespace("scales", quietly = TRUE)) install.packages("scales")
library(scales)

if (!requireNamespace("mgcv", quietly = TRUE)) install.packages("mgcv")
library(mgcv)

if (!requireNamespace("pso", quietly = TRUE)) install.packages("pso")
library(pso)
```

## AKKSAM

```{r AKSSAM, warning = FALSE, message=FALSE}
# AKSSAM
source("Required_Scripts/AKSSAM.R")
```

## GeDS

```{r GeDS, warning = FALSE, message=FALSE}
if (!requireNamespace("GeDS", quietly = TRUE)) install.packages("GeDS")
library(GeDS) 
```

# Data

Function to simulate the data given the sample size.

```{r Poisson Simulated Data 1}
simulate.data.pois1 = function(n){
  
  # This case study has m = 3 covariates
  m = 3
  
  n <- n                  # Nº instances
  X = matrix(1, n, m)     # Initialize the instance mat.
  FF = matrix(1, n, m)    # Initialize the func. values mat.
  f_total = rep(0, n)     # Linear estimator
  
  for (i in 1:m) {
    x <- runif(n, 0, 1)
    X[, i] <- x
    
    if (i %% m == 1) {
      f <- 1/2 * sin(4*pi*x)*log(20*x + 1)
    } else if (i %% m == 2) {
      f <- (x^3 + sin(2*pi*(x^7 - 2*x^3)))
    } else {
      f <- 1/2 * (sin(6*pi*x) + 3*x - 5* x^2 + cos(3*pi*x))
    }
    f_centered <- (f - mean(f))
    FF[, i] <- f_centered
    f_total <- f_total + f_centered
  }
  
  eta <- f_total            # Linear estimator
  lambda = exp(eta)         # Mean
  y = rpois(n, lambda)      # Response

  # Output
  list(X = X,
       y = y,
       FF = FF,
       eta = eta)
}
```

The simulated data consists on the additive model given by the following covariate effects

```{r Poisson Covariates}
#| fig-width: 10
#| fig-height: 3
#| fig-align: center
#| out-width: "100%"

# Simulate data
ll = simulate.data.pois1(n = 500)
X = ll$X
y = ll$y
FF = ll$FF

# Format the dataframe
df_long <- map2_dfr(
  as.data.frame(X), as.data.frame(FF),
  ~ tibble(X = .x, FF = .y),
  .id = "Covariate"
)

df_long <- df_long %>%
  mutate(Covariate = factor(Covariate, labels = paste("Covariate", 1:3)))

covariates = paste("Covariate", 1:3)

y_range <- range(df_long$FF)

# Build one plot per covariate
plots <- lapply(1:3, function(i) {
  df_i <- df_long[df_long$Covariate == covariates[i], ]
  
  ggplot(df_i, aes(x = X, y = FF)) +
    geom_line(color = "grey20", linewidth = 2) +
    scale_y_continuous(limits = y_range) +
    theme_light(base_size = 14) +
    labs(
      x = bquote(x),
      y = bquote(f[.(i)](x))
    ) +
    theme(
      panel.grid = element_line(color = "grey95")
    )
})

# Combine into a 1x3 grid
wrap_plots(plots, ncol = 3)
```


# Pipeline

## Fitting function

```{r}
fitting_function = function(X_train, X_test, y_train, y_test, eta_test,
                            ndx, m, bdeg, lambda.init,
                            maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                            grid){
  
  # Nº of instances and covariates
  n = length(y_train)
  family = "poisson"
  
  ## AKSSAM -------------------------------------------------------------------- 
  # Train the model
  time1 <- system.time({
    model1 <- AKSSAM(X_train, y_train, family, lambda.init, ndx, bdeg, 
                     maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, 
                     epsilon)
  })[3] %>% unname
  K_sel1 <- model1$K_sel
  alpha <- as.vector(model1$alpha.new)
  
  # Test
  Design_list = vector("list", m + 1)
  Design_list[[1]] = matrix(1, length(y_train), 1) 
  for (t in 1:m){Design_list[[t+1]] = my.bbase4(X_test[,t], K_sel1[[t]], bdeg[t])}
  B_star <- do.call(cbind, Design_list) 
  eta1 <- B_star %*% alpha
  predictions1 <- exp(eta1)
  
  # Grid
  Grid_list = vector("list", m)
  grid_basis_length = rep(0,m)
  for (t in 1:m){     
    Grid_list[[t]] = my.bbase4(grid[,t], K_sel1[[t]], bdeg[t])
    grid_basis_length[t] = dim(Grid_list[[t]])[2]
  }
  grid1 <- grid
  for (t in 1:m){
    B = Grid_list[[t]]
    if (t == 1){
      index_0 = 2
      index_1 = 1 + grid_basis_length[1]
    }else{
      index_0 = 2 + sum(grid_basis_length[1:(t-1)])
      index_1 = index_0 + grid_basis_length[t] - 1
    }
    grid1[ ,t] <- B %*% alpha[index_0:index_1]
  }
  
  # Metrics
  MSE1 <- mean((eta_test - eta1)^2)
  MAE1 <- mean(abs(eta_test - eta1))
  EDF1 <- Size1 <- ncol(B_star)
  
  loglik <- sum(y_test * log(predictions1) - predictions1 - lfactorial(y_test))
  AIC1 <- -2 * loglik + 2 * EDF1
  BIC1 <- -2 * loglik + log(n) * EDF1
  
  
  ## P-splines   ---------------------------------------------------------------
  # Train the model
  train <- data.frame(y = y_train, as.data.frame(X_train))
  terms <- paste0("s(V", 1:m, ", bs = 'ps', k = ", ndx + bdeg + 1, ")")             
  formula <- as.formula(paste("y ~", paste(terms, collapse = " + ")))
  time3 <- system.time({
    model3 <- mgcv::gam(formula, data = train, 
                        family = poisson(link = "log"), method = "REML")
  })[3] %>% unname
  
  # Test
  test = as.data.frame(X_test)
  eta3 <- predict(model3, newdata = test, type = "link")
  predictions3 <- exp(eta3)
  
  # Grid
  grid3 <- predict(model3, newdata = as.data.frame(grid), type = "terms")
  
  # Metrics
  MSE3 <- mean((eta_test - eta3)^2)
  MAE3 <- mean(abs(eta_test - eta3))
  EDF3 <- sum(model3$edf)
  Size3 <- length(model3$coefficients)
  
  
  loglik <- sum(y_test * log(predictions3) - predictions3 - lfactorial(y_test))
  AIC3 <- -2 * loglik + 2 * EDF3
  BIC3 <- -2 * loglik + log(n) * EDF3

  
  ## GeDS ----------------------------------------------------------------------
  terms_GeDS   <- paste0("f(V", 1:m, ")")
  formula_GeDS <- as.formula(paste("y ~", paste(terms_GeDS, collapse = " + ")))
  time4 <- system.time({
    model4 <- GeDS::NGeDSgam(formula_GeDS, family =  poisson(link = "log"), data = train, internal_knots = ndx[[1]] - 1)
  })[3] %>% unname
  K_sel4 <- model4$internal_knots$cubic.int.knots %>% unname
  
  
  # Test
  eta4 <- as.vector(predict(model4, n = 4, newdata = test, type = 'link'))
    predictions4 <- exp(eta4)
  
  # grid
  grid4 <- do.call(cbind, lapply(terms_GeDS, function(term) {
    matrix(
      predict(model4, n = 4, newdata = as.data.frame(grid), base_learner = term),
      ncol = 1, dimnames = list(NULL, term)
    )
  }))
  
  # Metrics
  MSE4  <- mean((eta_test - eta4)^2)
  MAE4  <- mean(abs(eta_test - eta4))
  EDF4 <- Size4 <- sum(sapply(knots(model4, n = 4L, options ="all"), length) %>% unname - bdeg - 1) + 1 # intercept
  loglik4          <- sum(y_test * log(predictions4) - predictions4 - lfactorial(y_test))
  AIC4             <- -2 * loglik4 + 2 * EDF4
  BIC4             <- -2 * loglik4 + log(n) * EDF4
  
  ## Assignment   --------------------------------------------------------------
  return(list(
    # AKSSAM algorithm results
    AKSSAM = list(MSE = MSE1,
                  MAE = MAE1,
                  Grid = grid1,
                  K_sel = K_sel1,
                  EDF = EDF1,
                  Size = Size1,
                  AIC = AIC1,
                  BIC = BIC1,
                  Time = time1),
    # P-splines algorithm results
    PSplines = list(MSE = MSE3,
                    MAE = MAE3,
                    Grid = grid3,
                    EDF = EDF3,
                    Size = Size3,
                    AIC = AIC3,
                    BIC = BIC3,
                    Time = time3),
    # GeDS algorithm results
    GeDS = list(MSE = MSE4,
                  MAE = MAE4,
                  Grid = grid4,
                  K_sel = K_sel4,
                  EDF = EDF4,
                  Size = Size4,
                  AIC = AIC4,
                  BIC = BIC4,
                  Time = time4)
  ))
}
```

## Covariates Plot

```{r}
covariates_plot <- function(Results, grid, N, m, n){
  
  ## True covariates's effects and X
  ll = simulate.data.pois1(n)
  X = ll$X
  FF = ll$FF
  
  ## Miscellanea
  methods <- c('AKSSAM', 'P-splines', 'GeDS')
  colors <- c("True effect" = "grey50", "AKSSAM" = "#F8766D", "P-splines" = "#00BFC4", 'GeDS' = "#EE9A49")
  
  ## Calculate the mean estimated covariates' effects over the simulations
  # AKSSAM algorithm
  Grid_AKSSAM_matrices <- lapply(Results, function(x) x$AKSSAM$Grid)
  mean_Grid_AKSSAM <- Reduce("+", Grid_AKSSAM_matrices) / N
  temp <- lapply(Results, function(x) x$AKSSAM$K_sel)
  K_sel_AKSSAM <- lapply(1:m, function(i){
    unlist(lapply(1:N, function(t){
      v <- temp[[t]][[i]]
      return(v[5:(length(v)-4)])
    }))
  })
  
  # P-Splines
  Grid_Psp_matrices <- lapply(Results, function(x) x$PSplines$Grid)
  mean_Grid_Psp <- Reduce("+", Grid_Psp_matrices) / N
  
  # GeDS algorithm
  Grid_GeDS_matrices <- lapply(Results, function(x) x$GeDS$Grid)
  mean_Grid_GeDS <- Reduce("+", Grid_GeDS_matrices) / N
  mean_Grid_GeDS <- apply(mean_Grid_GeDS, 2, function(x) x - mean(x)) ## Possible misspecification
  temp <- lapply(Results, function(x) x$GeDS$K_sel)
  K_sel_GeDS <- lapply(1:m, function(i){
    unlist(lapply(1:N, function(t){
      v <- temp[[t]][[i]]
      return(v)
    }))
  })
  
  # Combined grids
  mean_grids <- list(
    "AKSSAM" = mean_Grid_AKSSAM,
    "P-splines" = mean_Grid_Psp,
    'GeDS' = mean_Grid_GeDS
  )

  
  ## Marginalized covariates' effects
  plot_list <- vector('list', m)
  plot_list <- lapply(1:m, function(t) {
    
    
    plot_df <- do.call(rbind, lapply(methods, function(method) {
      data.frame(
        x = grid[, t],
        y = mean_grids[[method]][, t],
        Method = method
      )
    }))
    o <- order(X[, t])
    true_df <- data.frame(
      x = X[o, t],
      y = FF[o, t],
      Method = "True effect"
    )
    plot_df <- rbind(plot_df, true_df)
    plot_df$Method <- factor(plot_df$Method,
                         levels = c("P-splines", "AKSSAM", 'GeDS', "True effect"))


    y_range <- range(plot_df$y, na.rm = TRUE)
    y_min <- y_range[1]
    y_max <- y_range[2]
    y_base <- y_min - 0.4 * (y_max - y_min)
    dy <- (y_max - y_min) * 0.07
    
    rug_df_AKSSAM <- if (length(K_sel_AKSSAM[[t]]) > 0) {
        data.frame(x = K_sel_AKSSAM[[t]], 
                   ymin = y_base + dy, 
                   ymax = y_base + 2*dy)
      } else {
        data.frame(x = numeric(0), ymin = numeric(0), ymax = numeric(0))
      }
      
      rug_df_GeDS <- if (length(K_sel_GeDS[[t]]) > 0) {
        data.frame(x = K_sel_GeDS[[t]], 
                   ymin = y_base, 
                   ymax = y_base + dy)
      } else {
        data.frame(x = numeric(0), ymin = numeric(0), ymax = numeric(0))
      }
      
    
    plot <- ggplot(plot_df, aes(x = x, y = y, color = Method)) +
      geom_line(linewidth = 3, alpha = 0.6, data = subset(plot_df, Method == "True effect")) +
      geom_line(linewidth = 1.1, alpha = 0.8, data = subset(plot_df, Method != "True effect")) +
      # 'Density' of selected knots
      geom_segment(data = rug_df_AKSSAM,
                 aes(x = x, xend = x, y = ymin, yend = ymax),
                 color = "#F8766D", linewidth = 2, alpha = 2/N) +
       geom_segment(data = rug_df_GeDS,
                   aes(x = x, xend = x, y = ymin, yend = ymax),
                   inherit.aes = FALSE,
                   color = "#EE9A49", linewidth = 2, alpha = 2/N) +
      scale_color_manual(values = colors) +
      theme_minimal() +
      scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
      labs(title = paste0('Covariate ', t),
           color = 'Curve') +
      theme_light() + 
      theme(
        legend.position = "bottom",
        axis.title = element_blank(),
        panel.grid.minor = element_blank()
      )

      return(plot)
  })
  
  
  ## Combine the plots
  combined <- wrap_plots(plot_list, nrow = 1, guides = "collect") +
  plot_annotation(
    title = paste0("Poisson Simulation, n = ", n),
    theme = theme(
      plot.title = element_text(size = 15, face = "bold"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.background = element_rect(fill = alpha("white", 0.7), color = NA)
    )
  )
  
  ## Output
  return(combined)
}
```


## Error Plot

```{r}
error_plot <- function(Results){
  
  methods <- c("AKSSAM", "PSplines", "GeDS")
  metrics <- c("MSE", "MAE", "EDF", "Size", "Time", "BIC")
  metric_labels <- c("MSE", "MAE", "EDF", "Basis Size", "Time", "BIC")
  
  all_data <- list()
  for (method in methods) {
    for (metric in metrics) {
      values <- sapply(Results, function(x) x[[method]][[metric]])
      temp_df <- data.frame(
        Value = values,
        Metric = metric,
        Method = method
      )
      all_data[[paste0(method, "_", metric)]] <- temp_df
    }
  }
  df <- bind_rows(all_data)
  df$Metric <- recode_factor(df$Metric,
    MSE = "MSE",
    MAE = "MAE",
    EDF = "EDF",
    Size = "Basis Size",
    Time = "Time",
    BIC = "BIC"
  )
  df$Metric <- factor(df$Metric, levels = metric_labels)
  df$Metric <- factor(df$Metric, levels = c("MSE", "MAE", "Time", "EDF", "Basis Size", "BIC"))
  
  df$Method <- recode(df$Method,
    "PSplines" = "P-splines"
  )
  method_colors <- c(
    "AKSSAM"   = "#F8766D",
    "P-splines" = "#00BFC4",
    'GeDS' = "#EE9A49"
  )
  
  # Boxplot of all performance metrics
  Performance_plot <- ggplot(df, aes(x = Method, y = Value, color = Method)) +
    geom_boxplot(position = position_dodge(0.75), width = 0.6, fill = NA) +
    facet_wrap(~ Metric, scales = "free_y") +
    scale_color_manual(values = method_colors) +
    labs(title = "Performance Metrics",
         x = "Method", y = "Value", color = "Algorithm") +
    theme_light(base_size = 14) +
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      axis.text.x = element_text(size = 8),
    )

  
  ## Output
  return(Performance_plot)
}
```

## Wrapper

```{r}
wrapper_function <- function(N, n, m,
                             maxiter1, maxiter2, maxiter3, tol1, tol2, tol3, epsilon,
                             ndx, bdeg, lambda.init){
  
  current_env <- sys.frame(sys.nframe())
  list2env(as.list(current_env), envir = globalenv())
  
  # Set seed for reproducibility
  set.seed(1234)
  
  Results <- rep(list(
    list(
      AKSSAM = list(MSE = NA, MAE = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      PSplines = list(MSE = NA, MAE = NA, Grid = NA,
                      EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA),
      GeDS = list(MSE = NA, MAE = NA, Grid = NA, K_sel = vector("list", m),
                    EDF = NA, Size = NA, AIC = NA, BIC = NA, Time = NA)
    )
  ), N)
  
  ## Create evaluation grid
  grid <- sapply(1:m, function(j) seq(0, 1, length.out = 100))
  
  ## Loop which performs N iterations of the simulation
  i <- 1
  try_tolerance <- 3 # Tolerance for retries before aborting (2 tries)
  try_attempts <- 0  # Nº of attempts
  total_fails <- 0   # Total Nº of fails
  while (i <= N) {
    
    # Simulate once the situation with error handling
    try_result <- try({
      # Simulate the data
      data <- simulate.data.pois1(n) 
      X_train <- X_test <- data$X
      y_train <- y_test <- data$y
      eta_test <- data$eta
      # Fit the models
      result <- purrr::quietly(fitting_function)(
        X_train = X_train,
        X_test = X_test,
        y_train = y_train,
        y_test = y_test,
        eta_test = eta_test,
        ndx = ndx,
        m = m,
        bdeg = bdeg,
        lambda.init = lambda.init,
        maxiter1 = maxiter1,
        maxiter2 = maxiter2,
        maxiter3 = maxiter3,
        tol1 = tol1,
        tol2 = tol2,
        tol3 = tol3,
        epsilon = epsilon,
        grid = grid
      )$result
      
      # Store the result, increase index and restart try_attempts
      Results[[i]] <- result
      print(paste('Simulataion ', i, 'done.'))
      i <- i + 1
      try_attempts <- 0
    }, silent = FALSE)
    
    # Error handling
    if (inherits(try_result, "try-error")) {
      # Increase try_attempts
      try_attempts <- try_attempts + 1
      total_fails <- total_fails + 1
      print(paste('try_attempt:', try_attempts))
    }
    
    # Check if the attempt tolerance is met
    if (try_attempts == try_tolerance){
      warning('Simulation error: Attempt tolerance met \n')
      break
    }
  }
  
  
  ## Output
  return(Results)
}
```


## Processing Function

```{r process_function}
process_function <- function(Results, N, n){
  
  ## Miscelanea
  m <- 3
  grid <- sapply(1:m, function(j) seq(0, 1, length.out = 100)) %>% as.matrix
  
  ## Covariates estimated effect plot 
  cov_plot  <- covariates_plot(Results, grid, N, m, n)
  
  
  ## Error metrics plot and values
  err_plot <- error_plot(Results)
  
  methods <- c("AKSSAM", "PSplines", "GeDS")
  metrics <- c("MSE", "MAE", "AIC", "BIC", "EDF", "Size", "Time")
  
  err_metrics <- list()
  for (method in methods) {
    for (metric in metrics) {
      key_mean <- paste0("mean_", metric, "_", method)
      key_sd   <- paste0("sd_", metric, "_", method)
      
      values <- sapply(Results, function(x) {
        x[[method]][[metric]]
      })
      
      err_metrics[[key_mean]] <- mean(values, na.rm = TRUE)
      err_metrics[[key_sd]]   <- sd(values, na.rm = TRUE)
    }
  }
  
  ## Output
  return(list(covariates_plot = cov_plot,        # Covariates estimated effect plot 
              err_plot = err_plot,               # Error metrics plot
              err_metrics = err_metrics          # Error metrics
  ))
}

```


# Load Data

Load the .RData if the simulations are not run.

```{r load, eval = FALSE}
load("simulation2.RData")
```

# Simulations

To process the results, call for results_n$covariates_plot, results_n$err_plot or  results_n$err_metrics.

## n = 300

```{r Poisson n 300}
ll_Poisson_n300 <- wrapper_function(N = 100, n = 300, m = 3,
                       maxiter1 = 100, maxiter2 = 100, maxiter3 = 100, 
                       tol1 = 1e-5, tol2 = 1e-5, tol3 = 1e-5, epsilon = 1e-5,
                       ndx = rep(40, 3), bdeg = rep(3, 3), lambda.init = rep(10, 3))
```

```{r Processing n 300, eval = FALSE}
results_n300 <- process_function(Results = ll_Poisson_n300, N = 100, n = 300)
```


## n = 500

```{r Poisson n 500}
ll_Poisson_n500 <- wrapper_function(N = 100, n = 500, m = 3,
                       maxiter1 = 100, maxiter2 = 100, maxiter3 = 100, 
                       tol1 = 1e-5, tol2 = 1e-5, tol3 = 1e-5, epsilon = 1e-5,
                       ndx = rep(40, 3), bdeg = rep(3, 3), lambda.init = rep(10, 3))
```

```{r Processing n 500, eval = FALSE}
results_n500 <- process_function(Results = ll_Poisson_n500, N = 100, n = 500)
```


## n = 800

```{r Poisson n 800}
ll_Poisson_n800 <- wrapper_function(N = 100, n = 800, m = 3,
                       maxiter1 = 100, maxiter2 = 100, maxiter3 = 100, 
                       tol1 = 1e-5, tol2 = 1e-5, tol3 = 1e-5, epsilon = 1e-5,
                       ndx = rep(40, 3), bdeg = rep(3, 3), lambda.init = rep(10, 3))
```

```{r Processing n 800, eval = FALSE}
results_n800 <- process_function(Results = ll_Poisson_n800, N = 100, n = 800)
```




# Summary Table

## Full table

```{r Poisson Summary Table, eval = FALSE}
# Initialize the table
poisson_table = data.frame(
  n = c(300, 500, 800)
)

# Define indexes for n
n_vals <- c(300, 500, 800)

# Iterate through all possible n's
row_index <- 1
for (i in 1:3) {
  # Construct object name
  obj_name <- paste0("results_n", n_vals[i])
    
  # Fill the appropriate row
  poisson_table[row_index, names(results_n300$err_metrics)] <- get(obj_name)$err_metrics
    
  # Update the row index
  row_index <- row_index + 1
}



# Format the table such that the error terms are included as mean (sd)
poisson_table2 <- poisson_table

# Obtain all mean columns and iterate along them
mean_cols <- grep("^mean_", names(poisson_table), value = TRUE)
for (mean_col in mean_cols) {
  # Get the corresponding sd column
  sd_col <- sub("^mean_", "sd_", mean_col)
  # Format the column
  formatted_col <- sprintf("%.3f (%.3f)", poisson_table[[mean_col]], poisson_table[[sd_col]])
  poisson_table2[[mean_col]] <- formatted_col
}
# Erase sd columns
poisson_table2 <- poisson_table2[ , !grepl("^sd_", names(poisson_table2))]

names(poisson_table2) = c('n',
                          'MSE AKSSAM', 'MAE AKSSAM', 'AIC AKSSAM', 'BIC AKSSAM', 'EDF AKSSAM', 'Size AKSSAM', 'Time AKSSAM',
                          "MSE P-Splines", "MAE P-Splines", "AIC P-Splines", "BIC P-Splines", "EDF P-Splines", "Size P-Splines", 'Time P-splines')


# Display the Table
knitr::kable(poisson_table2, format = "html", caption = "Mean Results for the Poisson Scenario")

# LaTex code
xtable::xtable(poisson_table2)
```

## Reduced table

```{r, eval = FALSE}
reduced_poisson_table2 <- poisson_table2[, c(1,2,3,5,6,7,8,9,10,12,13,14,15)]

names(reduced_poisson_table2) = c('n', 
                           'MSE AKSSAM', 'MAE AKSSAM', 'BIC AKSSAM', 'EDF & Size AKSSAM',
                           "MSE P-Splines", "MAE P-Splines", "BIC P-Splines", "EDF P-Splines", "Size P-Splines")

# Display the Table
knitr::kable(reduced_poisson_table2, format = "html", caption = "Mean Results for the Poisson Scenario")


# LaTex code
xtable::xtable(reduced_poisson_table2)
```

# Save the RData

```{r}
# Checkpoint
save.image(file = "simulation3_VTest.RData")
```








